---
title: "Assignment 1"
author: "Michael LaVallee"
date: "9/18/2019"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### 1. Calculate the folowing sum
```{r}
s1 = c(1:2019)
sum(s1)
```

```{r}
s2 = c(1:2019)
sum(s2^3)
```

```{r}
s3<-c(1:2019)
sum(s3^s3)
```

```{r}
s4 = c(1:2019)
y = c(1,-1)
sum((s4*y)^s4)
```
```{r}
s5 = c(1:2019)
sum(1/(s5^2))

```
```{r}
s6 = c(1:2019)
sum(1/(s6))
```
```{r}
s7 = c(1:2019)
sum(1/(s7^3))
```
```{r}
s8 = c(1:2019)
y = c(1,-1)
sum(1/(s8*y))
```
#### 2. Generate random variable
###### a.
```{r}
a = rnorm(1000,mean=10,sd=1)
mean(a)
sd(a)
```
###### b.
```{r}
b = rnorm(1000,mean=10,sd=1)
a = b > 10
sum(a)
```
###### c.
```{r}
c = rnorm(1000,mean=10,sd=1)
hist(c)

```
###### d.
```{r}
d = rnorm(1000,mean=2,sd=1)
x = (d >1)
sum(x/1000)

```
#### 3. Consider an experiment of tossing a fair dice
```{r}
x = c(1:6)
y=sample(x,1000,replace=TRUE)
mean(y)
sd(y)
z=y>5
sum(z)
table(y)
prop.table(table(y))
plot(prop.table(table(y)))
```
#### 4. Consider an experiment of tossing a dice 3 times. 
```{r}
x = c(1:6)
y = matrix(sample(x,3000,replace=TRUE),nrow = 1000)
sum(y[,1]>y[,2]+y[,3])/1000
sum(y[,1]^2>y[,2]^2+y[,3]^2)/1000
```
#### 5. Using simulation, estimate the probability of getting three tails in a row when tossing a coin 3 times
```{r}
x = c(0,1)
y = matrix(sample(x,3000,replace = TRUE) , nrow = 1000)
sum(y[,1]+y[,2]+y[,3]==0)/1000

```
#### 6.Using simulation, estimate the probability of getting three tails in a row when tossing a coin 10 times
```{r}
x = c(0,1)
y = matrix(sample(x,30,replace = TRUE) , nrow = 10)
sum(y[,1]+y[,2]+y[,3]==0)/1000
```
#### 7 Central Limit Theorem
```{r}
x=runif(100000,0,1)
y=matrix(data=x,nrow = 100)
z=rowsum(y,c(1:100))
hist(z)


```
#### 7 titanic data
```{r}

d = read.csv("C:/Users/student/Documents/Senior Year/Fall Semester/R Analytics/Data Set/titanic.csv")
```
#### 8 Use knitr::kable function to nicely print out the first 10 rows of the data in markdown

```{r}

knitr::kable(d)
d[1:10,]
```
#### 9 Use is.na function and sum function to count the total number of missing values in the data.
```{r}
a=is.na(d)
sum(a)
colSums(a)

```
#### 10 Calculate the average Age of the passengers. You may want to use the parameter na.rm = TRUE in the function mean

```{r}
x <- c (d[,6])
mean(x,na.rm =TRUE)
```
#### 11 Replace the missing values of age by the average age calculated previously
```{r}
d[is.na(d)] <- 29.7

```
#### 12 Remove columns Name, PassengerID, Ticket, and Cabin.
```{r}
d$PassengerId = NULL
d$Name = NULL
d$Ticket = NULL
d$Cabin = NULL
```
#### 13 . Calculate the mean age of female passengers
```{r}
mean(d[d$Sex=='female',]$Age)

```
#### 14 Calculate the median fare of the passengers in Class 1
```{r}
median(d[d$Pclass == "1",]$Fare)
```
#### 15 Calculate the median fare of the female passengers that are not in Class 1
```{r}
median(d[d[d$Pclass != "1",]$Sex == "female",] $Fare)
```
#### 16. Calculate the median age of survived passengers who are female and Class 1 or Class 2,
```{r}
median(d[d[d[d$Pclass!="3",]$Sex == "female",]$Survived == "1",]$Age, na.rm = TRUE)
```
#### 17. Calculate the mean fare of female teenagers survived passengers
```{r}
mean(d[(d$Survived == "1") & (d$Age >=13)& (d$Age <=19) & (d$Sex == "female"),]$Fare)

```
#### 18. Calculate the mean fare of female teenagers survived passengers for each class
```{r}
one = mean(d[d[(d$Survived == "1") & (d$Age >=13)& (d$Age <=19) & (d$Sex == "female"),]$Pclass == "1",]$Fare)
one
two = mean(d[d[(d$Survived == "1") & (d$Age >=13)& (d$Age <=19) & (d$Sex == "female"),]$Pclass == "2",]$Fare)
two
three = mean(d[d[(d$Survived == "1") & (d$Age >=13)& (d$Age <=19) & (d$Sex == "female"),]$Pclass == "3",]$Fare)
three
```
#### 19. Calculate the ratio of Survived and not Survived for passengers who are who pays more than the average fare
```{r}
avgFare=mean(d$Fare)
survived=sum(d[d$Survived == "1",]$Fare > avgFare)
notsurvived = sum(d[d$Survived == "0",]$Fare > avgFare)

survived/notsurvived
```
#### 20. Add column that standardizes the fare (subtract the mean and divide by standard deviation) and name it sfare
```{r}
FARE_mean=mean(d$Fare)
sd_mean=sd(d$Fare)
Standard_Fare=(d$Fare-FARE_mean)/sd_mean
cbind(d,Standard_Fare)

```
#### 21 Add categorical variable named cfare that takes value cheap for passengers paying less the average fare and takes value expensive for passengers paying more than the average fare.
```{r}
Fare_Mean = mean(d$Fare)
value = d$Fare > Fare_Mean

value [value == "TRUE"] <- "Expensive" 
value [value == "FALSE"] <- "Cheap"

cbind(d,value)

```
#### 22 Add categorical variable named cage that takes value 0 for age 0-10, 1 for age 10-20, 2 for age 20-30,and so on
```{r}
cage = d$Age
cage[d$Age >=0] <- "0"
cage[d$Age >10] <- "1"
cage[d$Age >20] <- "2"
cage[d$Age >30] <- "3"
cage[d$Age >40] <- "4"
cage[d$Age >50] <- "5"
cage[d$Age >60] <- "6"
cage[d$Age >70] <- "7"
cage[d$Age >80] <- "8"


cbind(d,cage)


```
#### 23.  Show the frequency of Ports of Embarkation. It appears that there are two missing values in the Embarked variable. Assign the most frequent port to the missing ports. Hint: Use the levels function to modify the categories of categorical variables.
```{r}
table(d$Embarked)
levels(d$Embarked)= c("S","C","Q","S")
table(d$Embarked)

```






















